---
title: "RBridgeWeek3Assignment"
author: "Bin Lin"
date: "2016-7-31"
output: html_document
---

This week's assignment is to access the Mushrooms Dataset from UCI repository: https://archive.ics.uci.edu/ml/datasets/Mushroom. Then I will need to create a subset of this data with meaningful column names and row names. 

In order to gain access to the web data, I should install the RCurl package, then extract the data from the url and store it into a variable called raw_data which I am going to use later.
```{r}
install.packages("RCurl", repos = "http://cran.us.r-project.org")
library(RCurl)
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"
raw_data <- getURL(url)
```

The following code reads the data, and returns a data structure which is basically a data.frame.Setting stringAsFactors to False, can prevent character columns from being converted ti factor column. 
```{r}
mushroom_df <- read.csv(text = raw_data, header=FALSE, sep=",", stringsAsFactors = FALSE)
class(mushroom_df)
head(mushroom_df)
str(mushroom_df)
```


I am subsetting the data by choosing 4 columns from the data.frame.
```{r}
mushroom_subset <-subset(mushroom_df, select = c(V1, V4, V6, V22))
class(mushroom_subset)
head(mushroom_subset)
str(mushroom_subset)
```

I renamed the column names.
```{r}
colnames(mushroom_subset) <- c("class", "cap-color", "odor", "population")
head(mushroom_subset)
```


The raw data contains letters that are abbreviations of many attributes.I am replacing those abbraviations with some more meaningful words which can be found from the dataset description on the url. 
```{r}
mushroom_subset$"class"[mushroom_subset$"class"=="e"] <- "edible"
mushroom_subset$"class"[mushroom_subset$"class"=="p"] <- "poisonous"

mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="n"] <- "brown"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="b"] <- "buff"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="c"] <- "cinnamon"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="g"] <- "gray"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="r"] <- "green"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="p"] <- "pink"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="u"] <- "purple"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="e"] <- "red"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="w"] <- "white"
mushroom_subset$"cap-color"[mushroom_subset$"cap-color"=="y"] <- "yellow"

mushroom_subset$"odor"[mushroom_subset$"odor"=="a"] <- "almond"
mushroom_subset$"odor"[mushroom_subset$"odor"=="l"] <- "anise"
mushroom_subset$"odor"[mushroom_subset$"odor"=="c"] <- "creosote"
mushroom_subset$"odor"[mushroom_subset$"odor"=="y"] <- "fishy"
mushroom_subset$"odor"[mushroom_subset$"odor"=="f"] <- "foul"
mushroom_subset$"odor"[mushroom_subset$"odor"=="m"] <- "musty"
mushroom_subset$"odor"[mushroom_subset$"odor"=="n"] <- "none"
mushroom_subset$"odor"[mushroom_subset$"odor"=="p"] <- "pungent"
mushroom_subset$"odor"[mushroom_subset$"odor"=="s"] <- "spicy"

mushroom_subset$"population"[mushroom_subset$"population"=="a"] <- "abundant"
mushroom_subset$"population"[mushroom_subset$"population"=="c"] <- "clustered"
mushroom_subset$"population"[mushroom_subset$"population"=="n"] <- "numerous"
mushroom_subset$"population"[mushroom_subset$"population"=="s"] <- "scattered"
mushroom_subset$"population"[mushroom_subset$"population"=="v"] <- "several"
mushroom_subset$"population"[mushroom_subset$"population"=="y"] <- "solitary"
```

This is the final data.frame which concludes the data transformation. 
```{r}
head(mushroom_subset)
```